---
title: "Tidying-up-data-lesson"
author: "Cezanna"
date: "10/9/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyr)
library(dplyr)
```

to access 'filter' from the 'stats' package:
'stats::filter()'

# Read in and clean up data
```{r}
catch_original <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1", method = "libcurl"),
                    stringsAsFactors = FALSE)
head(catch_original)
```

To get pipeline operatore use command shift m. %>% 
Select the colums we want using `select()`

```{r}
catch_data <- catch_original %>% 
  #select(Region, Year, Chinook, Sockeye, Coho, Pink, Chum) you can do this, or the way below
  select(-All, -notesRegCode)

```

```{r}
summary(catch_data)
```

Change the values in the Chinook column to numeric using `mutate()`

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)

summary(catch_clean$Chinook)
```

investigate using `which()` and `is.na()`

```{r}
i <- which(is.na(catch_clean$Chinook))
i
```

401 is the number of the row where the n/a is

```{r}
catch_original[i,]
```

*use `mutate` to change the I to a 1
*use `mutate` to coerce Chinook column to numeric

```{r}
catch_clean <- catch_data %>% 
  mutate(Chinook = ifelse(Chinook %in% c("I", "l"), 1, Chinook)) %>% 
  mutate(Chinook = as.numeric(Chinook))

head(catch_clean)
```

# Tidy

Move from wide format to tall format using `pivot_(longer)`

```{r}
catch_long <- catch_clean %>% 
  pivot_longer(cols = -c(Region, Year),
               names_to = "species",
               values_to = "catch")
head(catch_long)
```

^ Not Pivoting over Region or Year so use -c

```{r}
catch_wide <- catch_long %>% 
  pivot_wider(names_from = species,
              values_from = catch)

head(catch_wide)
```

`rename()` catch to catch_thousands

```{r}
# catch_long <- catch_long %>% 
#   rename(catch_thousands = catch) %>% 
#   mutate(catch = catch_thousands * 1000) %>% 
#   select(-catch_thousands)

catch_long <- catch_long %>% 
  mutate(catch = catch * 1000)
  
  head(catch_long)
```


# Summarize

mean catch by region and species
```{r}
mean_region <- catch_long %>% 
  group_by(Region, species) %>% 
  summarise(mean_catch = mean(catch),
            n_obs = n())

mean_region


```

mean sockeye region

```{r}
mean_sockeye_region <- catch_long %>% 
  filter(species == "Sockeye") %>% 
  group_by(Region) %>% 
  summarise(mean_catch = mean(catch))

mean_sockeye_region
```

```{r}
annual_catch <- catch_long %>% 
  group_by(Year) %>% 
  summarise(total_catch = sum(catch)) %>% 
  arrange(total_catch)

annual_catch
```

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1",
                            method = "libcurl"),
                        stringsAsFactors = FALSE) %>% 
    select(code, mgmtArea)

head(region_defs)
```

```{r}
catch_joined <- left_join(catch_long, region_defs, by = c("Region" = "code"))

head(catch_joined)

```
```{r}
region_defs <- region_defs %>% 
  rename(Region = code)

catch_joined <- left_join(catch_long, region_defs, by = "Region")
catch_joined <- left_join(catch_long, region_defs)

```

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", remove = FALSE)
```
```{r}
dates_df %>% 
  separate(date, into = c("month", "day", "year"), sep = "/", remove = FALSE) %>% 
  unite(date_2, year, month, day, sep = "-")
```

