---
title: "My Wild Website"
author: "Cezanna"
date: "10/10/2019"
output:
  html_document: default
  pdf_document: default
---

```{r, warning = FALSE, message = FALSE}
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
library(leaflet)

```


```{r}
esc <- read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = F)
```

```{r}
annual_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-", remove = F)

head(annual_esc)
  
```


if you only want to show the plot and not the code, use {r, echo = FALSE}
if you want to hide the results {r, results = "hide"}
```{r, results = "hide"}
species_list <- c("Chinook", "Sockeye", "Chum", "Pink", "Coho")

annual_esc <- esc %>% 
  separate(sampleDate, c("Year", "Month", "Day"), sep = "-", remove = F) %>% 
  mutate(Year = as.numeric(Year)) %>% 
  group_by(Year, SASAP.Region, Species) %>% 
  summarise(total_escapement = sum(DailyCount)) %>% 
  filter(Species %in% c(species_list))



head(annual_esc)
```

# Visualize Escapement
aes stands for aesthetics! this has to do with the structure of the plot, not the actual aesthetics of the plot. LOL. 


Total Escapement by Region and Speciea
```{r}
ggplot(data = annual_esc, mapping = aes(x = Species, y = total_escapement, fill = SASAP.Region)) + 
  geom_col()
```



Kodiak Escapement Over Time by Species
```{r}

my_theme <- theme_bw() +
  theme(legend.position = "bottom")
  
  
ggplot(data = filter(annual_esc, SASAP.Region == "Kodiak"),
       mapping = aes(x = Year, y = total_escapement, color = Species)) +
  geom_point() +
  geom_line() +
  scale_color_brewer(palette = "BuPu" ) +
  ylab("Escapement (no. fish)") +
  ggtitle( "Kodiak Salmon Escapement") +
  my_theme


```
 
 Escapement over time: all regions
 
```{r, fig.width = 7, fig.height = 12, echo = FALSE}
fig1 <- ggplot(annual_esc, aes(x = Year, y = total_escapement, color = Species)) +
  geom_point() +
  geom_line () + 
  scale_color_brewer(palette = "RdPu" ) +
  facet_wrap(~SASAP.Region, scales = "free_y", ncol = 1) +
  my_theme

fig1

ggsave("escapement.png", fig1, width = 12, height = 7, units = "in")
```

# Map Sampling Locations

Create a table with unique sampling lcoations, latitudes, and longitudes

```{r}
locations <- esc %>% 
  distinct(Location, Latitude, Longitude) %>% 
  drop_na

head(locations)
```

```{r}
datatable(locations)
```

```{r}
leaflet(locations) %>% 
  addTiles() %>% 
  addMarkers(lng = ~ Longitude, lat = ~ Latitude, popup = ~ Location)
```

